\textsf{Are large language models superhuman chemists?}

### 2024-03-28 src/tex/ms.tex
update overviedw
edits mar 28

\clearpage
\begin{abstract}
    Large language models (LLMs) have found widespread interest as they can perform tasks they have not been explicitly trained on. 
    This is relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in the form of text.
    LLMs have shown promise to address these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.

    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be required to improve models and mitigate potential harms. 
    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, against the expertise of human chemists.

    We curated more than 7000 question-answer pairs about a wide array of subfields of the chemical sciences, evaluated leading models and found that the best models, on average, outperform the best human chemists in our study. 
    The models, however, struggle with some chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, e.g., about the safety profile of chemicals.

@@

\clearpage
\begin{refsection}
\section{Introduction}
\Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination.\autocite{kung2023performance} 
They also have been shown to design and autonomously perform chemical reactions when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4, Boiko_2023, bran2023chemcrow}
While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
In any case, the promise of these models is that they show the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
This led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}

Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}

In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science.
For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
Many of these insights are in form of text in scientific publications. 
Thus, operating on such texts might be our best way of unlocking these insights and to learn from them.
This might ultimately lead to general copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
Such a usage mode is in particular interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous}

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used in the inverse way to predict toxic ones. 
Yet, it is important to realize that, while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools have been available for years.
Still, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
However, even for experts the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa} have been proposed, none of them have been widely accepted.
This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.

@@

To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specialization.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, including safety-related ones, might be very misleading.
We find that there are still plenty of limitations for current models to overcome, to find use in autonomous systems for chemists. 
Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.  %and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.

@@

Importantly, our large pool of questions encompasses a wide range of topics.
This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different subfields of the chemical sciences (see \Cref{sec:meth-topic} for details on how we assigned topics).
The distribution of topics is also evident from \Cref{fig:question_diversity} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
In this representation, semantically similar questions are close to each other, and we color the points based on a classification into \variable{output/num_topics.txt} topics. 
It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as a large distinct cluster.

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/question_count_barplot.pdf}
    \caption{\textbf{Number of questions for different topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on word-embeddings of the questions. 
    The figure shows that not all aspects of chemistry are equally represented in our corpus. The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}. This figure represents the combined count of \glstext{mcq} and open-ended questions.}
    \label{fig:topic_barplot}
    \script{plot_statistics.py}
\end{figure}

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} To obtain this figure, we embedded questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results). We then project the embeddings into a two-dimensional space using \gls{pca}. We color the points based on a classification into topics. Safety related aspects cover a large part of the figure that is not covered by questions from other topics.}
    \label{fig:question_diversity}
    \script{plot_question_diversity.py}
\end{figure}

While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).

\paragraph{\enquote{Tiny} subset}
It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process).
We also used this subset to seed the web application for the human baseline study.

@@

\paragraph{Benchmark suite design} Because the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a special way.
For instance, the Galactica model\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
Current benchmarking suites, however, do not account for such special treatment of scientific information.
To address this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
For instance, molecules represented in \gls{smiles} are enclosed in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
This allows the model to treat the \gls{smiles} string different from other text. 
\chembench can seamlessly handle such special treatment in an easily extensible way because the questions are stored in an annotated format.

Since many widely utilized systems do only provide access to text completions (and not the raw model outputs), \chembench is designed to operate on text completions.
This is also important with the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the final outputs of the system that would also be used in a real-world application. 
Hence, we use them for our evaluations.

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance. 
    Since the humans did not answer all the questions, this plot is based on the subset of questions that most humans answered.
    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, partially correct answers are also considered as incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often not find the correct answer within the specified number of calls.
    In this case, we consider the answer as incorrect. 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

\paragraph{System performance} 
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study in this overall metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
Many other models also outperform the average human performance. 
It is, however, interesting to note that the Galactica model that was trained specifically for scientific applications is outperformed by more general models.

Given the large interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is striking. 
Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
With the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), however, this often did not allow the system to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
This observation highlights the importance of communicating and measuring not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified all questions in the \chembench corpus into these topics based on hand-crafted rules as well as classifier models operating on the question texts.
We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correct by the models.
    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that was answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}
In this spider chart, the worst score for every dimension is zero (no question answered correctly) and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule). 
These findings also shine an interesting light on the value of textbook-inspired questions. 
A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

@@

It is important to note that the model performance for some topics, however, is underestimated in the current evaluation. 
This is because models that are provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
For instance, models might refuse to provide answers about cyanides. 
To overcome this, direct access to the model weights would be required, and we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
This is facilitated with the tooling \chembench provides, thanks to which new models can be automatically added by contributors in an open science fashion.

@@

% Estimate of difficulty
\paragraph{Confidence estimates} One might wonder whether the models are capable of estimating if they will be able to answer a question correctly. 
If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.

@@

\subparagraph{Tool augmented models}
In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
For this we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
In addition, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
We implemented the systems using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.

@@

\paragraph{Participants}
Users were open to report about their experience in chemistry. 
Overall, \variable{output/num_users_with_education_info.txt} did so. 
Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.

>>>>

\clearpage
\begin{abstract}
    Large language models (LLMs) have gathered widespread interest due to their abilities to process human language and perform tasks they have not been explicitly trained on. 
    This is particularly relevant for the chemical sciences, which face the problem of small and diverse datasets that are frequently in diverse forms of text.
    LLMs have shown promise to address these issues and are increasingly being harnessed to predict chemical properties, optimize reactions, and even design and conduct experiments autonomously.

    However, we still have only very limited systematic understanding of the chemical reasoning capabilities of LLMs, which would be helpful in understanding limitations, create more performant models and mitigate potential harms. 
    Here, we introduce \enquote{\chembench,} an automated framework designed to rigorously evaluate the chemical knowledge and reasoning abilities of state-of-the-art LLMs, against the expertise of human chemists.

    We curated more than 7000 question-answer pairs for a wide array of subfields of the chemical sciences, evaluated leading open and closed-source LLMs and found that the best models, on average, outperform the best human chemists in our study. 
    The best-performing models, however, struggle with certain chemical reasoning tasks that are easy for human experts and provide overconfident misleading predictions, such as describing the safety profile of chemicals.

@@

\clearpage
\begin{refsection}
\section{Introduction}
\Glspl{llm} are \gls{ml} models trained on massive amounts of text to complete sentences. 
Aggressive scaling of these models has led to a rapid increase in their capabilities,\autocite{brown2020language,zhong2024benchmarking} with the leading models now being able to pass in some evaluations the United States Medical Licensing Examination.\autocite{kung2023performance} 
They also have been shown to design and autonomously perform chemical reactions when augmented with external tools such as web search and synthesis planners.\autocite{openai2024gpt4, Boiko_2023, bran2023chemcrow}
While some see \enquote{sparks of \gls{agi}} in them,\autocite{bubeck2023sparks} others consider them as \enquote{stochastic parrots}---i.e., systems that only regurgitate what they have been trained on.\autocite{bender2021dangers}
Nevertheless, the promise of these models is that they have shown the ability to solve a wide variety of tasks they have not been explicitly trained on.\autocite{bommasani2021opportunities, anderljung2023frontier}
This has led to tremendous economic interest and investment in such generative models, with an expected market of more than \$1.3 trillion (almost \$30 billion for drug discovery applications) by 2032.\autocite{bloomberg}

Chemists and materials scientists have quickly caught on the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language}.\autocite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lála2023paperqa, caufield2023structured, gupta2022discomat} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}

In addition, since a lot---if not most---of the information about chemistry is currently stored and communicated in text there is a strong reason to believe that there is still a lot of untapped potential in \glspl{llm} for chemistry and materials science \autocite{miret2024llms}.
For instance, most insights in chemical research do not directly originate from data stored in databases but rather from the scientists and their ability to interpret data. 
Many of these insights are in form of text in scientific publications. 
Thus, operating on such texts might be our best way of unlocking these insights and to learn from them.
This might ultimately lead to general copilot systems for chemists that can provide answers to questions or even suggest new experiments based on vastly more information than a human could ever read.
Such a usage mode is in particular interesting in the face of recent advances in autonomous laboratories.\autocite{Boiko_2023, bran2023chemcrow, darvish2024organa, granda2018controlling, Angello_2022, coley2019robotic, Burger_2020, seifrid2022autonomous, sim2023chemos}

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used in the inverse way to predict toxic ones. 
Yet, it is important to recognize that, while \glspl{llm} can facilitate access to information and tools, this information about controlled chemicals and task-specific tools have been available for years.
Moreover, it is important to realize that the user base of such models is not only limited to experts in chemistry and materials science who critically reflect on every output such models produce. 
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes. 
However, even for experts the chemical understanding and reasoning capabilities are important as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from anecdotal reports, there is little evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

While some benchmarks based on university entrance exams\autocite{Zaki_2024, arora2023llms} or automatic text mining\autocite{song2023honeybee, wei2021chemistryqa, song-etal-2023-matsci} have been proposed, none of them have been widely adopted.
This is likely because they cannot automatically be used with black box (or tool-augmented) systems, do not cover a wide range of topics, or are not carefully validated by experts.
On top of that, the existing benchmarks are not designed to be used with models that support special treatment of molecules or equations and do not provide insights on how the models compare relative to experts.

@@

To contextualize the scores, we also surveyed more than \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specialization.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but in many cases, including safety-related ones, might be very misleading.
We find that there are still numerous limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists. 
Moreover, we conclude that carefully created broad benchmarks represent an important stepping stone for progress in this field.  %and underline the importance of careful evaluation of the capabilities of \glspl{llm} in the chemical sciences.

@@

Importantly, our large pool of questions encompasses a wide range of topics.
This can be seen, for example, in \Cref{fig:topic_barplot} in which we compare the number of questions in different subfields of the chemical sciences (see \Cref{sec:meth-topic} for details on how we assigned topics).
The distribution of topics is also evident from \Cref{fig:question_diversity} in which we visualize the questions in a two-dimensional space using a \gls{pca} on the embeddings of the questions.
In this representation, semantically similar questions are close to each other, and we color the points based on a classification into \variable{output/num_topics.txt} topics. 
It is clear that a focus of \chembench (by design) lies on safety-related aspects which in \Cref{fig:question_diversity} appear as a large distinct cluster across the embedding space.

\begin{figure}[!htb]
    \centering
    \includegraphics{figures/question_count_barplot.pdf}
    \caption{\textbf{Number of questions for different chemistry topics.} The topics have been assigned using a combination of a rule-based system (mostly based on the source the question has been sampled from) as well as a classifier operating on word-embeddings of the questions. 
    The figure shows that not all aspects of chemistry are equally represented in our corpus. The \chembench corpus, by design, currently focuses on safety-related aspects, which is also evident in \Cref{fig:question_diversity}. This figure represents the combined count of multiple choice questions (\glstext{mcq}) and open-ended questions.}
    \label{fig:topic_barplot}
    \script{plot_statistics.py}
\end{figure}

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\textwidth]{figures/question_diversity.pdf}
    \caption{\textbf{Principal component projection of embeddings of questions in the \chembench corpus.} This figure represents a projection of text embeddings from \chembench questions and answers using the BART model\autocite{bart} (using other embeddings, such as of OpenAI's ada model leads to qualitatively similar results). We project the embeddings into a two-dimensional space using \gls{pca} and color the points based on a classification into topics. Safety related aspects, the primary focus of \chembench, cover a large part of the figure that is not covered by questions from other topics.}
    \label{fig:question_diversity}
    \script{plot_question_diversity.py}
\end{figure}

While many of the existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of today's chemistry education and research practices.
For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions).

\paragraph{\enquote{Tiny} subset}
It is important to note that for routine evaluations, a smaller subset of the corpus might be more practical.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark. 
To address this, we also provide a subset (\variable{output/num_tiny_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus in which topics are more balanced than in the complete corpus (see \Cref{sec:subset-selection} for details on the curation process). This \enquote{tiny} also provided the seed of the web application for our human baseline study.

@@

\paragraph{Benchmark suite design} Given that the text used in scientific settings is different from typical natural language, many models have been developed that deal with such text in a targeted manner.
For instance, the Galactica model\autocite{taylor2022galactica} uses special tokenization or encoding procedures for molecules and equations. 
Current benchmarking suites, however, do not account for such special treatment of scientific information.
To address this, \chembench is designed to encode the semantic meaning of various parts of the question or answer.  
For instance, molecules represented in \gls{smiles} are enclosed in \texttt{[START\_SMILES][\textbackslash END\_SMILES]} tags. 
This allows the model to treat the \gls{smiles} string different from other text. 
\chembench can seamlessly handle such special treatment in an easily extensible way because the questions are stored in an annotated format.

Since many widely utilized systems only provide access to text completions (and not the raw model outputs), \chembench is also designed to operate on text completions.
This is also important given the growing number of tool-augmented systems that are deemed essential for building chemical copilot systems. 
Such systems can augment the capabilities of \glspl{llm} through the use of external tools such as search \glspl{api} or code executors.\autocite{schick2024toolformer, karpas2022mrkl, yao2022react}
In those cases, the \gls{llm} that returns the probabilities for various tokens (that are often used for model evaluations\autocite{Fourrier_Habib_Launay_Wolf}) is only a part of the whole system, and it is not clear how to interpret the probabilities in the context of the whole system.
The text completions, however, are the final outputs of the system that would also be used in a real-world application prompting us to them for our evaluations.

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered completely correctly. Horizontal bars indicate the performance of various models and highlight statistics of the human performance. 
    Since many of the humans did not answer all the questions, we created this plot based on the common subset of questions that most humans answered.
    The evaluation we use here is very strict as it only considers a question answered completely correctly or completely incorrectly, with partially correct answers being considered as incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    Systems with \enquote{ReAct} in the name are tool augmented, i.e., they can call external tools such as web search or Python code executors to better answer the questions.
    However, we limit those systems to a maximum of ten calls to the \gls{llm}. This constraint led the systems to often fail to find the correct answer within the specified number of calls. If the \gls{llm} exceeds ten calls, we consider the answer as incorrect. 
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

\paragraph{System performance} 
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar}. 
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study which we obtained via a custom web application (\url{chembench.org}) used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, Claude 3, outperforms the best human in our study according to the overall performance metric and vastly, by more than a factor of two, exceeds the average performance of the experts in our study.
Furthermore, many commercial models also outperform the average human performance. 
Interestingly, the Galactica model, which was trained specifically for scientific applications, underperforms compared to many advanced commerical and some open-source models and barely exceeds the random baseline.

Given the large interest in tool-augmented systems, the mediocre performance of these systems (GPT-3.5 and Claude 2 augmented with tools) in our benchmark is particuarly salient. 
Their lack of performance, however, is partially due to the fact that we limited the system to a maximum of ten \gls{llm} calls.
Using the default tool augmented setup (using the so-called ReAct method\autocite{yao2023react}), these systems were often unable to identify the correct answer (e.g., because it repeatedly tried to search for the answer on the web and did not find a solution within ten calls to the \gls{llm}).
This observation highlights the importance of communicating and measuring not only in predictive performance but also computational cost (e.g., in terms of \gls{api} calls) for tool-augmented systems.

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and classified all questions in the \chembench corpus into these topics based on hand-crafted rules as well as classifier models operating on the question texts.
We then computed the percentage of questions that the models, or humans, answered completely correctly for each topic.
\begin{figure}[!h]
    \centering
    \includegraphics{figures/all_questions_models_completely_correct_radar_human.pdf}
    \caption{\textbf{Performance of the models and humans on the different topics of the \enquote{tiny} subset of the \chembench corpus.} The radar plot shows the performance of the models and humans on the different topics of \enquote{tiny} subset of the \chembench corpus. The performance is measured as the fraction of questions that were answered completely correct by each of the models.
    The best score for every dimension is one (all questions answered correctly) and the worst score is zero (no question answered completely correctly). A larger colored area indicates a better performance.
    This figure shows the performance on the subset of questions that was answered by humans. The performance of models on the entire corpus is shown in \Cref{fig:performance_per_topic}.
    }
    \label{fig:all_questions_models_completely_correct_radar_human}
    \script{analyze_model_reports.py}
\end{figure}
In this spider chart, the worst score for every dimension is zero (no question answered correctly) and the best score is one (all questions answered correctly). 
Thus, a larger colored area indicates a better performance. 
One can observe that this performance varies widely across models and topics. 
While macromolecular chemistry and biochemistry receive relatively high scores for many models, this is not the case for topics such as chemical safety or analytical chemistry.
In the subfield of analytical chemistry the prediction of the number of signals observable in a \gls{nmr} spectrum proved difficult for the models (e.g., \variable{output/subset_scores/is_number_nmr_peaks_gpt4.txt} percent correct answers for GPT-4) while this question appeared easier (\variable{output/human_subset_scores/is_number_nmr_peaks.txt} percent correct) for trained humans.
Importantly, the human experts are given a drawing of the compounds, whereas models are only shown the \gls{smiles} string of a compound and have to use this to reason about the symmetry of the compound (i.e., to identify the number of diasterotopically distinct protons, which requires \emph{reasoning} about the topology and structure of a molecule). 
These findings also shine an interesting light on the value of textbook-inspired questions. 
A subset of the questions in the \chembench are based on textbooks targeted at undergraduate students. 
On those questions the models tend to perform better than on some of our semi-automatically constructed tasks (see \Cref{fig:performance_per_topic}).
For instance, while the overall performance in the chemical safety topic is low, the models would pass the certification exam according to the German Chemical Prohibition Ordinance based on a subset of questions we sampled from the corresponding question bank (e.g., \variable{output/subset_scores/is_gfk_gpt4.txt}\% correct answers for GPT-4, \variable{output/subset_scores/is_gfk_claude3.txt}\% for Claude 3, and \variable{output/human_subset_scores/is_gfk.txt}\% for the human experts).
While those findings are impacted by the subset of questions we sampled, the results still highlight that good performance on such question bank or textbook questions does not necessarily translate to good performance on other questions that require more reasoning.

@@

It is important to note that the model performance for some topics, however, is underestimated in the current evaluation procedure. 
This is because models that are provided via \glspl{api} typically have safety mechanisms that prevent them from providing answers that the provider deems unsafe.
For instance, models might refuse to provide answers about cyanides. 
To overcome this, direct access to the model weights would be required, and as such we strive to collaborate with the developers of frontier models to overcome this limitation in the future.
We aim to faciliate this with the tooling \chembench provides, thanks to which new models can be automatically added by contributors in an open science fashion.

@@

% Estimate of difficulty
\paragraph{Confidence estimates} One interesting question to investigate is whether the models are capable of estimating if they will be able to answer a question correctly. 
If they were able to do so, incorrect answers would be less problematic as one would be able to detect when an answer is incorrect.
To investigate this, we prompted\autocite{xiong2023llms} some of the top-performing models to estimate, on an ordinal scale, their confidence in their ability to answer the question correctly.

@@

\subparagraph{Tool augmented models}
In addition to directly prompting \glspl{llm}, we also investigated the performance of tool-augmented systems.
For this we investigated the 7B parameter \enquote{online} model of Perplexity.AI. 
Additionally, we utilized gpt-3.5-turbo as well as Claude 2, with ReAct-style tool augmentation.\autocite{yao2023react}
The latter two models had access to WolframAlpha, the ArXiv \gls{api}, a Python interpreter, as well as web search (using DuckDuckGo).
We implemented all systems using Langchain with the default prompts and constrained the system to a maximum of ten \gls{llm} calls.

@@

\paragraph{Participants}
Users were open to report about their experience in chemistry. 
Overall, \variable{output/num_users_with_education_info.txt} out of YYYY did so. 
Out of those, \variable{output/num_human_phd.txt} reported to have been awarded a Ph.D.
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree, and \variable{output/num_human_bachelor.txt} have a bachelor's degree.
